---
title: "Chemical EDA"
author: "Suheng Yao"
date: "2024-10-19"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(PubChemR)
```

```{r}
# Read in the cleaned dataset and select records of California and Florida
df <- read.csv("strawberry_cleaned.csv")
df <- df %>%
  select(-1) %>%
  filter(State == "CALIFORNIA" | State == "FLORIDA")
```

```{r}
# Split the California data and Florida data into two dataframes
df_cal <- df %>%
  filter(State == "CALIFORNIA")

df_flo <- df %>%
  filter(State == "FLORIDA")

# Find the distinct chemicals used in California and Florida
length(unique(df_cal$Chemical_Info))
length(unique(df_flo$Chemical_Info))

# Find out chemicals used by California but not used by Florida
diff_chem1 <- setdiff(unique(df_cal$Chemical_Info), unique(df_flo$Chemical_Info))
print(diff_chem1)

diff_chem2 <- setdiff(unique(df_flo$Chemical_Info), unique(df_cal$Chemical_Info))
print(diff_chem2)
```
After finding the distinct chemicals in each state, California used 19 more types of chemicals than Florida, and there are 58 types of chemicals used in California but not in Florida, on the other hand, there are 39 types of chemicals used in Florida but not in California. We can first explore those 58 chemicals used only in California:
```{r}
# Select records with those 58 chemicals
df_chem_cal <- df_cal %>%
  filter(Chemical_Info %in% diff_chem1)

# Create a new column to record the frequency of each chemical used
chem_freq <- df_chem_cal %>%
  count(Chemical_Info) %>%
  arrange(desc(n))

ggplot(chem_freq[1: 10,], aes(x=reorder(Chemical_Info, n), y=n))+
  geom_bar(stat = "identity")+
  theme_minimal()+
  coord_flip()+
  labs(x = "Chemical", y = "Frequency", 
       title = "Top 10 Chemicals Used Only in California")+
  theme(plot.title.position = "plot")
```
Plot above shows the top 10 most used chemicals in strawberry harvesting, we can further explore the properties of those 10 chemicals:
```{r message=FALSE, warning=FALSE}
# Find the index of GHS classification
GHS_searcher<-function(result_json_object){
  # check if the chemicals in the database first
  if (is.null(result_json_object) || 
      is.null(result_json_object[["result"]]) || 
      is.null(result_json_object[["result"]][["Hierarchies"]]) ||
      is.null(result_json_object[["result"]][["Hierarchies"]][["Hierarchy"]])){
    return("did not find the chemical in the database")
  }
  
  result<-result_json_object
  for (i in 1:length(result[["result"]][["Hierarchies"]][["Hierarchy"]])){
    if(result[["result"]][["Hierarchies"]][["Hierarchy"]][[i]][["SourceName"]]=="GHS Classification (UNECE)"){
      return(i)
    }
          
  }
}

# Use the output from GHS_searcher to access the hazard information
hazards_retriever<-function(index,result_json_object){
  
  # Check if GHS_searcher did not find the index
  if (is.character(index) && index == "did not find the chemical in the database") {
    return(index)
  }
  
  result<-result_json_object
  hierarchy<-result[["result"]][["Hierarchies"]][["Hierarchy"]][[index]]
  i<-1
  output_list<-rep(NA,length(hierarchy[["Node"]]))
  while(str_detect(hierarchy[["Node"]][[i]][["Information"]][["Name"]],"H") & i<length(hierarchy[["Node"]])){
    output_list[i]<-hierarchy[["Node"]][[i]][["Information"]][["Name"]]
    i<-i+1
  }
  return(output_list[!is.na(output_list)])
}

# Find the chemical information for the top 10 chemicals
chem_vec <- chem_freq[1: 10, "Chemical_Info"]
access_hazard <- function(chemical){
  results <- list()
  for (chem in chemical){
      result <- get_pug_rest(identifier = chem, 
                             namespace = "name", 
                             domain = "compound",
                             operation="classification", 
                             output = "JSON")
      ghs_index <- GHS_searcher(result)
      hazards <- hazards_retriever(ghs_index, result)
      results[[chem]] <- list(
        chemical_name = chem,
        chemical_hazards = ifelse(hazards == "did not find the chemical in the database", character(0), hazards)
      )
    
  }
  
  return(results)
}
  

hazard_info <- access_hazard(chem_vec)
hazards_df <- do.call(rbind, lapply(hazard_info, function(x) {
  data.frame(
    chemical_name = x$chemical_name,
    hazards = paste(x$chemical_hazards, collapse = ";"),
    stringsAsFactors = FALSE
  )
}))
knitr::kable(hazards_df, escape = FALSE)
```
