---
title: "Chemical EDA"
author: "Suheng Yao"
date: "2024-10-19"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
library(ggplot2)
library(tidyverse)
library(PubChemR)
library(kableExtra)
library(flextable)
```

```{r}
# Read in the cleaned dataset and select records of California and Florida
df <- read.csv("strawberry_cleaned.csv")

# Find states that used chemicals in growing strawberry
states_with_chemical_info <- df %>%
  filter(!is.na(Chemical_Info)) %>%
  distinct(State) %>%
  pull(State)

print(states_with_chemical_info)

```
Only California and Florida used chemicals when growing strawberries, we can filter out only those two states and compare the frequency of chemical used:
```{r, warning=FALSE}
df1 <- df %>%
  select(-1) %>%
  filter(State == "CALIFORNIA" | State == "FLORIDA")

# Draw a bar plot comparing the frequency
chem_state <- df1 %>%
  group_by(State) %>%
  summarise(Total_obs=n()) %>%
  arrange(desc(Total_obs))
print(chem_state)

ggplot(chem_state, aes(x = State, y = Total_obs, fill = State)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Total_obs), vjust = -0.3, size = 4) +
  theme_minimal() +
  labs(title = "Total Counts of Chemical Usage in Strawberry Harvesting",
       x = "State",
       y = "Total Times Chemicals Used") +
  theme(legend.position = "none")  # Remove legend
```
The frequency that California used chemicals is almost twice as the frequency of Florida.

```{r}
# Split the California data and Florida data into two dataframes
df_cal <- df %>%
  filter(State == "CALIFORNIA")

df_flo <- df %>%
  filter(State == "FLORIDA")

# Find the distinct chemicals used in California and Florida
chem_comparison <- data.frame(
  State = c("CALIFORNIA", "FLORIDA"),
  Distinct_chemicals = 
    c(length(unique(df_cal$Chemical_Info)),
      length(unique(df_flo$Chemical_Info)))
)


ggplot(chem_comparison, aes(x = State, y = Distinct_chemicals, fill = State)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = Distinct_chemicals), vjust = -0.3, size = 4) +
  theme_minimal() +
  labs(title = "Total Types of Chemicals Used in Strawberry Harvesting",
       x = "State",
       y = "Total Types of Chemicals Used") +
  theme(legend.position = "none")  # Remove legend
```
After finding the distinct chemicals in each state, California used 19 more types of chemicals than Florida, and there are 58 types of chemicals used in California but not in Florida:
```{r}
# Find out chemicals used by California but not used by Florida
diff_chem1 <- setdiff(unique(df_cal$Chemical_Info), unique(df_flo$Chemical_Info))
print(diff_chem1)
```

on the other hand, there are 39 types of chemicals used in Florida but not in California:
```{r}
diff_chem2 <- setdiff(unique(df_flo$Chemical_Info), unique(df_cal$Chemical_Info))
print(diff_chem2)
```

We can first explore top chemicals used in California:
```{r, warning=FALSE, message=FALSE}

# Create a new column to record the frequency of each chemical used
chem_measure <- df_cal %>%
  group_by(Measurement) %>%
  filter(Measurement == "MEASURED IN LB")

# Plot the top chemicals used in California
top_chem <- chem_measure %>%
  filter(!is.na(Value)) %>%
  group_by(Chemical_Info, Domain) %>%
  summarise(Value = sum(Value, na.rm = TRUE)) %>%
  arrange(desc(Value))

ggplot(top_chem[1:10,], 
       aes(x=reorder(Chemical_Info, Value), y=Value))+
  geom_bar(stat = "identity")+
  theme_minimal()+
  coord_flip()+
  labs(x = "Chemical", y = "Usage", 
       title = "Top 10 Chemicals Measured in LB in California")+
  theme(plot.title.position = "plot")
```
```{r}
domain_chem <- chem_measure %>%
  filter(!is.na(Value)) %>%
  group_by(Domain) %>%
  summarise(count = n()) %>%
  arrange(desc(count))
print(domain_chem)
```

Plot above shows the most used chemicals in LB in strawberry harvesting, we can further explore the properties of those chemicals:
```{r, message=FALSE, warning=FALSE}
# Find the index of GHS classification
GHS_searcher<-function(result_json_object){
  # check if the chemicals in the database first
  if (is.null(result_json_object) || 
      is.null(result_json_object[["result"]]) || 
      is.null(result_json_object[["result"]][["Hierarchies"]]) ||
      is.null(result_json_object[["result"]][["Hierarchies"]][["Hierarchy"]])){
    return("did not find the chemical in the database")
  }
  
  result<-result_json_object
  for (i in 1:length(result[["result"]][["Hierarchies"]][["Hierarchy"]])){
    if(result[["result"]][["Hierarchies"]][["Hierarchy"]][[i]][["SourceName"]]=="GHS Classification (UNECE)"){
      return(i)
    }
          
  }
}

# Use the output from GHS_searcher to access the hazard information
hazards_retriever<-function(index,result_json_object){
  
  # Check if GHS_searcher did not find the index
  if (is.character(index) && index == "did not find the chemical in the database") {
    return(index)
  }
  
  result<-result_json_object
  hierarchy<-result[["result"]][["Hierarchies"]][["Hierarchy"]][[index]]
  i<-1
  output_list<-rep(NA,length(hierarchy[["Node"]]))
  while(str_detect(hierarchy[["Node"]][[i]][["Information"]][["Name"]],"H") & i<length(hierarchy[["Node"]])){
    output_list[i]<-hierarchy[["Node"]][[i]][["Information"]][["Name"]]
    i<-i+1
  }
  return(output_list[!is.na(output_list)])
}

# Find the chemical information for the top chemicals
access_hazard <- function(chemical){
  results <- list()
  for (chem in chemical){
      result <- get_pug_rest(identifier = chem, 
                             namespace = "name", 
                             domain = "compound",
                             operation="classification", 
                             output = "JSON")
      ghs_index <- GHS_searcher(result)
      hazards <- hazards_retriever(ghs_index, result)
      results[[chem]] <- list(
        chemical_name = chem,
        chemical_hazards = ifelse(hazards == "did not find the chemical in the database", character(0), hazards)
      )
    
  }
  
  return(results)
}
  


```
```{r, message=FALSE, warning=FALSE}
chem_vec <- top_chem[1:10,]$Chemical_Info
hazard_info <- access_hazard(chem_vec)
hazards_df <- do.call(rbind, lapply(hazard_info, function(x) {
  data.frame(
    chemical_name = x$chemical_name,
    hazards = paste(x$chemical_hazards, collapse = ";"),
    stringsAsFactors = FALSE
  )
}))


hazards_df %>%
  flextable() %>%
  theme_vanilla() %>%
  fontsize(size = 10) %>%
  width(j = "chemical_name", width = 2.5) %>%  # Set width for first column
  width(j = "hazards", width = 5) %>%  # Set width for second column
  align(align = "left", part = "all") %>%
  set_table_properties(layout = "autofit")
```

Through the chemical information presented in the above table, seven out of the top ten chemicals have hazards information. Especially for QUINOLINE, which can cause damage to organs and toxic in contact. Also, most of those chemicals are very toxic to aquatic life, which will harm the sustainable development of the environment. Additionally, most of the top used chemicals belong to insecticide group, and this suggests that many of the chemicals used in California are used for pest control. This phenomenon could be the result of large production of strawberry and warm weather in California.

Now, we can further analyze the top chemicals used in strawberry growing in Florida:
```{r}
# Select records with Florida only chemicals
df_measure_lb <- df_flo %>%
  filter(Measurement=="MEASURED IN LB") %>%
  filter(!is.na(Value)) %>%
  filter(Chemical_Info != "TOTAL")

print(unique(df_measure_lb$Chemical_Info))
```
Ignoring "Total" variable, here are the eight chemicals that are used most in Florida when measurement in LB.

```{r, message=FALSE}
# Create a new column to record the frequency of each chemical used
top_chem <- df_measure_lb %>%
  group_by(Chemical_Info, Domain) %>%
  summarise(Value = sum(Value, na.rm = TRUE)) %>%
  arrange(desc(Value))

# Plot the top 10 chemicals used in California
ggplot(top_chem, aes(x=reorder(Chemical_Info, Value), y=Value))+
  geom_bar(stat = "identity")+
  theme_minimal()+
  coord_flip()+
  labs(x = "Chemicals", y = "Measured in LB", 
       title = "Top Chemicals Used in Florida")+
  theme(plot.title.position = "plot")
```
```{r, message=FALSE, warning=FALSE}
chem_vec_flo <- top_chem$Chemical_Info
hazard_info_flo <- access_hazard(chem_vec_flo)

hazards_df_flo <- do.call(rbind, lapply(hazard_info_flo, function(x) {
  data.frame(
    chemical_name = x$chemical_name,
    hazards = paste(x$chemical_hazards, collapse = ";"),
    stringsAsFactors = FALSE
  )
}))

hazards_df_flo %>%
  flextable() %>%
  theme_vanilla() %>%
  fontsize(size = 10) %>%
  width(j = "chemical_name", width = 2.5) %>%  # Set width for first column
  width(j = "hazards", width = 5) %>%  # Set width for second column
  align(align = "left", part = "all") %>%
  set_table_properties(layout = "autofit")
```
Through the table shown above, five out of top eight chemicals used in Florida have hazard information. The most toxic chemical is the azoxystrobin, which can be toxic if inhaled, and it also causes damage to organs with single exposure. Additionally, it has toxic effects to aquatic lives, which is bad for the environment. Also, 7 out of top 8 chemicals used are insecticide, this may relate to the location of Florida, which is already in the tropical area. The warmer weather brings more insects, so the farmers there need to use more insecticide for pest control. The principal pests on strawberry in Florida are the twospotted spider mite and lepidopterous larvae.

In summary, by comparing the chemicals used in California and Florida, California used more types of chemicals than Florida, also, the counts of chemicals used in growing strawberry were also higher for California, which is almost two times of Florida's value. However, California and Florida both used chemicals in insecticide that are toxic to the aquatic lives, environment and human's health, and it is important to suggest to the goverment to restrict the amount of harmful chemicals used in insecticides and pesticides, and they should be changed to more environmental friendly substitutes.
