---
title: "Strawberries Dataset EDA"
author: Suheng Yao
date: 2024 Sept 30
format: pdf
editor: visual
---

## Read and Explore the Data

```{r}
#| label: load libraries and set options
#| warning: false
#| message: false
#| 

library(knitr)  
library(kableExtra)
library(tidyverse)
library(ggplot2)
library(tidyr)
library(stringr)

```

We first read in the data and have an overview of the data:

```{r}
#| label: read data - glimpse
#| message: false 

strawberry <- read_csv("strawberries25_v3.csv", col_names = TRUE)

glimpse(strawberry)
```

We got 12699 rows and 21 columns.

## Data Cleaning

We can check for NA values and the columns with exact same values:
```{r}
# Identify columns with only NA values and remove them
na_num <- colSums(is.na(strawberry))
print(na_num)
strawberry <- strawberry[, na_num < nrow(strawberry)]
```
Since Week Ending, Zip Code, Region and Watershed contain only NA values, those four columns have to be removed.

```{r}
# Identify columns with exact same values and remove them
col_same <- sapply(strawberry, function(x) length(unique(x)) == 1)
print(col_same)
strawberry <- strawberry[, !col_same]
```
Since watershed_code and Commodity are the columns with the same values, they also have to be removed.

```{r}
glimpse(strawberry)
```
Now, there are 12669 rows and 15 variables.

Since the "State ANSI", "County ANSI", "Value" and "CV (%)" are all numeric values, but they were treated as character values in this data, it is necessary to convert them into numeric value:
```{r, warning=FALSE}
# convert two columns to numeric
strawberry$`State ANSI` <- as.numeric(strawberry$`State ANSI`)
strawberry$`County ANSI` <- as.numeric(strawberry$`County ANSI`)
strawberry$Value <- as.numeric(strawberry$Value)
strawberry$`CV (%)` <- as.numeric(strawberry$`CV (%)`)
```

Since the "Geo Level", "Ag District", "State" and "County" are all good indicators of the group, it is better to convert them into factor data type:
```{r}
# convert four columns to factor type
cols_convert <- c("Geo Level", "State", "Ag District")
strawberry[cols_convert] <- lapply(strawberry[cols_convert], as.factor)
```


### Examine The Structure of the Data

```{r}
#| label: explore organization 1 


## Check if every row is associated with a state

state_all <- strawberry |> distinct(State)

state_all1 <- strawberry |> group_by(State) |> count()

if(sum(state_all1$n) == dim(strawberry)[1]){print("Yes every row in the data is associated with a state.")}

rm(state_all, state_all1)

```

### Separate Composite Columns

Since variable "Data Item" contains too much information, we plan to separate the information in this column into a few new columns:

```{r}
#| label: split Data Item

# Step1: Split the Data Item variable first based on commas
strawberry <- strawberry %>%
  separate(`Data Item`, into = c("Fruit", "Category1", "Category2"), 
           sep = ",", fill = "right", extra = "merge")

# Step2: Split the Fruit column based on hyphen
strawberry <- strawberry %>%
  separate(Fruit, into = c("Fruit", "Operation"), 
           sep = "-", fill = "right", extra = "merge")

# Step3: Separate Category1 column based on hyphen
strawberry <- strawberry %>%
  separate(Category1, into = c("Product Type", "Product Details"), 
           sep = "-", fill = "right", extra = "merge")

# Step4: Separate Product Type column into Product Type and Measurement
strawberry <- strawberry %>%
  mutate(Measurement = ifelse(str_detect(`Product Type`, "MEASURED IN"), `Product Type`, NA_character_),
    `Product Type` = ifelse(str_detect(`Product Type`, "MEASURED IN"), NA_character_, `Product Type`))

# Step5: Separate Category2 column based on hyphen
strawberry <- strawberry %>%
  separate(Category2, into = c("Product Type1", "Product Details1"), 
           sep = "-", fill = "right", extra = "merge")

# Step6: Split the Product Type1 column to create another new column Measurement1
strawberry <- strawberry %>%
  separate(`Product Type1`, into = c("Product Type1", "Measurement1"), 
           sep = ",", fill = "right", extra = "merge")

# Step7: Put the measurement information in Product Type1 into Measurement column
strawberry <- strawberry %>%
  mutate(Measurement = ifelse(str_detect(`Product Type1`, "MEASURED IN| AVG"), `Product Type1`, Measurement),
    `Product Type1` = ifelse(str_detect(`Product Type1`, "MEASURED IN| AVG"), NA_character_, `Product Type1`))

# Step8: Split the Product Details1 column based on comma
strawberry <- strawberry %>%
  separate(`Product Details1`, into = c("Product Details1", "Measurement2"), 
           sep = ",", fill = "right", extra = "merge")

# Step9: Combine measurement2, measurement1 and measurement
strawberry <- strawberry %>%
  mutate(
    Measurement = coalesce(Measurement, Measurement1, Measurement2)
  ) %>%
  select(-Measurement1, -Measurement2)

# Step10: Combine Product Type and Product Type1, Product Details and Product Details1
strawberry <- strawberry %>%
  mutate(
    `Product Type` = coalesce(`Product Type`, `Product Type1`)
  ) %>%
  select(-`Product Type1`)

strawberry <- strawberry %>%
  mutate(
    `Product Details` = coalesce(`Product Details`, `Product Details1`)
  ) %>%
  select(-`Product Details1`)

# Step11: Trim the leading space
strawberry$Fruit <- str_trim(strawberry$Fruit, side = "both")
strawberry$Operation <- str_trim(strawberry$Operation, side = "both")
strawberry$`Product Type` <- str_trim(strawberry$`Product Type`, side = "both")
strawberry$`Product Details` <- str_trim(strawberry$`Product Details`, side = "both")
strawberry$Measurement <- str_trim(strawberry$Measurement, side = "both")
```
The "Data Item" column has been splitted into "Fruit", "Operation", "Product Type", "Product Details" and "Measurement".


Now, since the Domain and Domain Category columns have redundant and complex information, we consider split and combine the information in those two columns:
```{r}
unique(strawberry$Category2)
```

```{r}
#| label: Split and Transform Domain Category Variable and Domain Variable

# Step1: Split the Domain Category based on comma
strawberry <- strawberry %>%
  separate(`Domain Category`, into = c("Domain1", "Category1"), 
           sep = ",", fill = "right", extra = "merge")

# Step2: Focus on Domain1 column first, split it based on colon
strawberry <- strawberry %>%
  separate(Domain1, into = c("Domain2", "Category2"), 
           sep = ":", fill = "right", extra = "merge")

# Step3: Combine the Domain and Domain2 column
strawberry <- strawberry %>%
  mutate(
    Domain = coalesce(Domain, Domain2)
  ) %>%
  select(-Domain2)

# Step4: Split Category2 column

```

